Rising from the Ashes: How the Milky Way Got Its Scars

### 2024-04-29 paper/src/res.tex
More changes

\section{Results}\label{sec:results}
\subsection{Orbital Summary}\label{ssec:orbit}

\subsection{Abundance Distribution}\label{ssec:abundplane}
In Figure~\ref{fig:fig1}, briefly discussed in Section~\ref{sec:intro}, we show the abundance distribution of the Milky Way as well as two of our idealized merger simulations in the upper panels. A number of our idealized simulations exhibit either a bimodal or unimodal abundance distribution, and so we have selected two representative examples. In detail, the simulation labeled bimodal had orbital parameters of $R_0=142\,\kpc$, $V_0=116\,\kms$, and $\eta=0.4$. The unimodal simulation is identical, except that $R_0=129s\,\kpc$.

There are, of course, differences between the bimodal simulation and the Milky Way. First, the scaling in \MgFe{} is different -- in the simulation, the low-$\alpha$ sequence lies at $\sim0.2$, while in the Milky Way it is at about $\MgFe\sim0$. Second, in the Milky Way the high-$\alpha$ sequence neatly joins the low-$\alpha$ sequence at $\FeH\sim0$, while in the simulation the two actually diverge more at higher \FeH{}. Finally, at high \FeH{} the distribution bends upwards while in the simulation it bends downwards. The explanations for these differences are numerous, and are not the focus of the current work.

Finally, in the lower panels of Figure~\ref{fig:fig1}, we show the distributions of \MgFe{} at different fixed \FeH{}. The (blue, orange, red, green) lines show the \MgFe{} distribution at a \FeH{} of ($-0.5$, $-0.375$, $-0.25$, $0.25$), in bins of width $0.05\,\dex$. The distributions of \MgFe{} are offset so that they do not overlap. Here, the bimodality seen in the Milky Way is quite striking at lower metallicities. The peaks are well-separated, by $\sim0.2\,\dex$. In the bimodal simulation, the distribution is still clearly bimodal, but the peaks are less well-separated, by $\sim0.1\,\dex$. In the unimodal simulation, the distributions are clearly unimodal, with the hint of some structure in the $\FeH=0.25\,\dex$ bin.

\subsection{Build up of the Abundance Plane}\label{ssec:abundplane_build}
Next, we examine the build up of the abundance plane. First, in Figure~\ref{fig:before_after}, we show the abundance plane at various points along the merger. We have divided the history of the merger into ``before'' ($t < 1.5\,\Gyr$), ``during'' ($1.5\,\Gyr < t < 2.5\,\Gyr$), and ``after'' ($t > 2.5\,\Gyr$). These times have been chosen somewhat by the orbital history and somewhat by visually inspecting the movie of the simulation, and so these deliniations are, to an extent, arbitrary. Nonetheless, it is clear that the high-$\alpha$ sequence forms before and during the merger (middle panels), and that the low-$\alpha$ sequence forms after the merger (right panel). We show a dashed line given by $-0.1\FeH + 0.31$, which is a crude way of separating the high- and low-$\alpha$ sequence.

@@

As expected, we see a period of suppressed star formation in the bimodal simulation, which is also preceded by a period of enhanced star formation. The dip in SFR is not as pronounced as it was when we plotted the SFH at a particular iron (Figure~\ref{fig:before_after_sfh}). The period of a suppressed SFR is associated with a sudden drop in \MgFe{}. In $\sim250\,\Myr$, \MgFe{} drops from $\sim0.35$ to $\sim0.2\,\dex$. Once the SFR recovers $\sim500\,\Myr$ later, \MgFe{} again rises, but only to $\sim0.28\,\dex$. Roughly speaking, when star formation ceases, \MgFe{} has a value of $\sim0.35\,\dex$. When it recovers, \MgFe{} has a value of $\sim0.28\,\dex$.

This behavior is present in the unimodal simulation (right panel) -- it does not display a period of suppressed star formation, and there is no associated drop in \MgFe{}. It does still have a period of enhanced star formation, but this is actually coincident with an enhancement in \MgFe{}.

@@

>>>>

\section{Results}\label{sec:results}
\subsection{Abundance Distribution}\label{ssec:abundplane}
In Figure~\ref{fig:fig1}, briefly discussed in Section~\ref{sec:intro}, we show the abundance distribution of the Milky Way as well as two of our idealized merger simulations in the upper panels. A number of our idealized simulations exhibit either a bimodal or unimodal abundance distribution, and so we have selected two representative examples. In detail, the simulation labeled bimodal had orbital parameters of $R_0=142\,\kpc$, $V_0=116\,\kms$, and $\eta=0.4$. The unimodal simulation is identical, except that $R_0=129\,\kpc$.

There are, of course, differences between the bimodal simulation and the Milky Way. First, the scaling in \MgFe{} is different -- in the simulation, the low-$\alpha$ sequence lies at $\sim0.2$, while in the Milky Way it is at about $\MgFe\sim0$. Second, in the Milky Way the high-$\alpha$ sequence neatly joins the low-$\alpha$ sequence at $\FeH\sim0$, while in the simulation the two actually diverge more at higher \FeH{}. Finally, at high \FeH{} the distribution bends upwards while in the simulation it bends downwards.

Finally, in the lower panels of Figure~\ref{fig:fig1}, we show the distributions of \MgFe{} at different fixed \FeH{}. The (blue, orange, red, green) lines show the \MgFe{} distribution at a \FeH{} of ($-0.5$, $-0.375$, $-0.25$, $0.25$), in bins of width $0.05\,\dex$. The distributions of \MgFe{} are offset (but not rescaled) so that they do not overlap. Here, the bimodality seen in the Milky Way is quite striking at lower metallicities. The peaks are well-separated, by $\sim0.2\,\dex$. In the bimodal simulation, the distribution is still clearly bimodal, but the peaks are less well-separated, by $\sim0.1\,\dex$. In the unimodal simulation, the distributions are clearly unimodal, with the hint of some structure in the $\FeH=0.25\,\dex$ bin.

\subsection{Build up of the Abundance Plane}\label{ssec:abundplane_build}
Next, we examine the build up of the abundance plane. First, in Figure~\ref{fig:before_after}, we show the abundance plane at various points along the merger. We have divided the history of the merger into ``before'' ($t < 1.5\,\Gyr$), ``during'' ($1.5\,\Gyr < t < 2.5\,\Gyr$), and ``after'' ($t > 2.5\,\Gyr$). These times have been chosen somewhat by the orbital history and somewhat by visually inspecting the simulation, and so these deliniations are, to an extent, arbitrary. Nonetheless, it is clear that the high-$\alpha$ sequence forms before and during the merger (middle panels), and that the low-$\alpha$ sequence forms after the merger (right panel). We show a dashed line given by $-0.1\FeH + 0.31$, which is a crude way of separating the high- and low-$\alpha$ sequence.

@@

The merger appears to induce a starburst, where the SFR rises from $\sim3.5\,\Msunyr$ at $\sim1\Gyr$ to a peak of $\sim12\,\Msunyr$ at $\sim2\,\Gyr$. This starburst is then followed by a quiescent period, with a depth of $\sim1\,\Msunyr$ and a duration of $\sim500\,\Myr$. The dip in SFR is not as pronounced as it was when we plotted the SFH at a particular iron (Figure~\ref{fig:before_after_sfh}). This quiescent period is associated with a sudden drop in \MgFe{}. In $\sim250\,\Myr$, \MgFe{} drops from $\sim0.35$ to $\sim0.2\,\dex$. Once the SFR recovers $\sim500\,\Myr$ later, \MgFe{} again rises, but only to $\sim0.28\,\dex$.

This behavior is present in the unimodal simulation (right panel) -- it does not display a period of suppressed star formation, and there is no associated drop in \MgFe{}. It does still have a period of enhanced star formation, but this is actually coincident with an enhancement in \MgFe{}. This sequence of events is further emphasized in Figure~\ref{fig:alpha_vs_tform}, which shows a transparent scatter plot of \MgFe{} as a function of formation time. In the bimodal simulation, there is a clear gap in \MgFe{} during the period of quiescence followed by the brief formation of $\alpha$-deficient stars. In the unimodal simulation, on the other hand, there is no gap and the brief formation of $\alpha$-enhanced stars.

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{alpha_vs_tform.pdf}
  \caption{The formation times and \MgFe{} values of star particles in the bimodal (blue) and unimodal (orange) simulations. In the bimodal simulation, a gap in stellar ages is briefly followed by the formation of $\alpha$-deficient star particles. In the unimodal simulation, no gap is present and there is briefly the formation of $\alpha$-enhanced star particles at the same time. The \MgFe{} values of the unimodal simulation have been offset by $-0.3$ for clarity.}
  \label{fig:alpha_vs_tform}
\end{figure}
